---
title: "Lobyerev V1"
author: "Itai van Rijn"
date: "February 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## Implimatation of Lobyerev & Hoffman (2008) Selectivity model

This document describes the implimatation of the selectivity algorithem described in Lobyrev, Feodor, and Matthew J. Hoffman. "A morphological and geometric method for estimating the selectivity of gill nets." Reviews in Fish Biology and Fisheries 28.4 (2018): 909-924.

###Step 1: Import field data on fish catch and gill net properties

####Step 1.1: Import catch data

Catch data structure:
Column 1: 'Mesh_size' - name of the net (later conected to net properties under the same name)
Column 2: 'Length_group' - Fish total length (cm)
Column 3: 'Wedged' - Number of wedged individuals 
Column 4: 'Tangled' - Number of tangled individuals
```{r}
Catch.data.Cod <- read.csv("~/kinneret modeling/selectivity/R code Pyodor paper/Data/Catch data Cod.csv")
#Show the first 6 lines:
print(head(Catch.data.Cod))
```

####Step 1.2: Import table of net properties
Net properties data structure: (See Figure 2 in paper)
Column 1: 'Mesh_size' - mesh size (knot to knot) (mm)
Column 2: 'y' - Smaller of the angels between mesh threads

```{r}
net.properties <- read.csv("~/kinneret modeling/selectivity/R code Pyodor paper/Data/net properties.csv")
#Show table:
print(net.properties)
```

####Step 1.3: Add angel in Radian
```{r}
net.properties$Radian=0.018*net.properties$y
#Show table:
print(net.properties)
```

####Step 1.4: Define the angel between the end of upper and lower jaws (phi)
```{r}
phi_deg=7
phi_radian=0.018*phi_deg
```

####Step 1.5: Calculate the jaw length approximated by the linear function
```{r}
slope_jaw=1.02
intecept_jaw=3.41
Catch.data.Cod$Jaw_length=intecept_jaw + (slope_jaw*Catch.data.Cod$Length_group)
#round Jaw_length to the nearest 0.1
#Catch.data.Cod$Jaw_length=round(Catch.data.Cod$Jaw_length,digits = 1)
#Calculate h
Catch.data.Cod$h=Catch.data.Cod$Jaw_length * sin(phi_radian) * 2
#round h to the nearest 0.1
#Catch.data.Cod$h=round(Catch.data.Cod$h,digits = 1)
print(head(Catch.data.Cod))
```
####Step 1.6: Merge the catch data with the net data
```{r}
Catch.data.Cod=merge(Catch.data.Cod,net.properties,by="Mesh_size")
#print first 6 lines
print(head(Catch.data.Cod))
```

###Step 2: Calculate P(O|C) and P(Th|C)
For calculating P(O|C) use eq. 4 in the paper

```{r}
#Calculate sin,cos,tan
Catch.data.Cod$sin=round(sin(0.5*Catch.data.Cod$Radian),digit=1)
Catch.data.Cod$cos=round(cos(0.5*Catch.data.Cod$Radian),digit=1)
Catch.data.Cod$tan=round(tan(0.5*Catch.data.Cod$Radian),digit=1)
#Calculate P(O|C) Eq. 4
Catch.data.Cod$POC=(((Catch.data.Cod$cos * Catch.data.Cod$Mesh_size)-Catch.data.Cod$h)*((Catch.data.Cod$sin*Catch.data.Cod$Mesh_size)-(Catch.data.Cod$tan*Catch.data.Cod$h)))/(Catch.data.Cod$cos*Catch.data.Cod$sin*(Catch.data.Cod$Mesh_size)^2)
Catch.data.Cod$POC=round(Catch.data.Cod$POC,digits = 2)
#Calculate P(Th|c)
Catch.data.Cod$PThC=1-Catch.data.Cod$POC
#Print first 6 rows
print(head(Catch.data.Cod))
```

###Step 3: Calculate P(W|E)
!!!!I just import the data from yur Excel, but do not know how the calculation was made
```{r}
PWE <- read.csv("~/kinneret modeling/selectivity/R code Pyodor paper/Data/PWE.csv", stringsAsFactors=FALSE)
#merge with table 'Catch.data.Cod'
Catch.data.Cod=merge(Catch.data.Cod,PWE ,by=c("Mesh_size",    "Length_group"))
```

###Step 4: Calculate Eq. 2

####Step 4.1: Calculate P(E|O) by the linear equation

```{r}
##Get the min and max size groups for each net
#Create table
net_PEO=data.frame(Mesh_size=unique(Catch.data.Cod$Mesh_size))
#Length group interval
Length_group_interval=2
#The minimal size for each net
min_wedged=data.frame(Catch.data.Cod %>%
   filter(!is.na(Wedged)) %>%
  group_by(Mesh_size) %>%
  summarize(min_size = min(Length_group, na.rm = TRUE)))
#The maximal size for each net
max_wedged=data.frame(Catch.data.Cod %>%
   filter(!is.na(Wedged)) %>%
  group_by(Mesh_size) %>%
  summarize(max_size = max(Length_group, na.rm = TRUE)))
#merge
net_PEO=merge(net_PEO,min_wedged,by="Mesh_size")
net_PEO=merge(net_PEO,max_wedged,by="Mesh_size")
#Substract and add the Length group interval
net_PEO$min_size=net_PEO$min_size-(Length_group_interval/2)
net_PEO$max_size=net_PEO$max_size+(Length_group_interval/2)
#Calculate the linear function
net_PEO$slope=-1/(net_PEO$max_size-net_PEO$min_size)
net_PEO$intercept=1-(net_PEO$slope*net_PEO$min_size)
#print
print(net_PEO)
#Calculate PEO
#Merge slope and intercept to data
Catch.data.Cod=merge(Catch.data.Cod,net_PEO,by="Mesh_size")
Catch.data.Cod$PEO=Catch.data.Cod$intercept+(Catch.data.Cod$Length_group)*Catch.data.Cod$slope
Catch.data.Cod$PEO=round(Catch.data.Cod$PEO,digits = 2)
#Remove PEO values if no fish were wedged
Catch.data.Cod[is.na(Catch.data.Cod$Wedged),"PEO"]=NA
#Print first 6 lines
print(head(Catch.data.Cod))
```

####Step 4.2: Claculate Eq.2
```{r}
Catch.data.Cod$Ntotal=Catch.data.Cod$Wedged/(Catch.data.Cod$POC*Catch.data.Cod$PWE*Catch.data.Cod$PEO)
#Print first 6 lines
print(head(Catch.data.Cod))
```

###Step 5: Calculate CPUE
(table 21)
```{r}
#Define the number of trials
n_trials=11
#Aggregate catch for each mesh size
Catch.data.Cod$CPUE=(Catch.data.Cod$Wedged+Catch.data.Cod$Tangled)/n_trials
#Print first 6 lines
print(head(Catch.data.Cod))
```

###Step 6: Calculate Nw per hour
(table 23)
```{r}
#Define number of hours of single (??) field trial
n_hours=12
Catch.data.Cod$Ntotal_per_hour=Catch.data.Cod$Ntotal/n_hours
#Print first 6 lines
print(head(Catch.data.Cod))
```

###Step 7: 
####Step 7.1: SLl,t for each length class as Ntotal-Qt
(table 24)
```{r}
Catch.data.Cod$SL_l_t=Catch.data.Cod$Ntotal_per_hour-Catch.data.Cod$CPUE
#Print first 6 lines
print(head(Catch.data.Cod))
```

####Step 7.2: Nlim table
(table 22c)
```{r}
###The next data frame will be an experimental input
n_lim=data.frame(Mesh_size=c(20,25,30),Nlim=c(36,32,12))
#Sum CPUE per net
CPUE=data.frame(Catch.data.Cod %>%
  group_by(Mesh_size) %>%
  summarize(CPUE = sum(CPUE, na.rm = TRUE)))
#merge to n_lim table
n_lim=merge(n_lim,CPUE,by="Mesh_size")
#Calculate tau (table 25)
n_lim$tau=n_hours/(-log(1-(n_lim$CPUE/n_lim$Nlim)))
n_lim$N_AP=((n_lim$Nlim*(1.71))/n_lim$tau)*n_hours
```


