---
title: "Lobyerev V1"
author: "Itai van Rijn"
date: "February 24, 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)
```

## implementation of Lobyerev & Hoffman (2018) Selectivity model

This document describes the implementation of the selectivity algorithm described in Lobyrev, Feodor, and Matthew J. Hoffman. "A morphological and geometric method for estimating the selectivity of gill nets." Reviews in Fish Biology and Fisheries 28.4 (2018): 909-924.

###Step 1: Import field data on fish catch and gill net properties

####Step 1.1: Import catch data

Catch data structure:
Column 1: 'Mesh_size' - name of the net (later connected to net properties under the same name)
Column 2: 'Length_group' - Fish total length (cm)
Column 3: 'Wedged' - Number of wedged individuals 
Column 4: 'Tangled' - Number of tangled individuals
```{r}
Catch.data.Cod <- read.csv("~/kinneret modeling/selectivity/R code Feodor paper/Lobyerev-selectivity/Data/Catch data Cod.csv")
#Show the first 6 lines:
print(head(Catch.data.Cod))
```

####Step 1.2: Import table of net properties
Net properties data structure: (See Figure 2 in paper)
Column 1: 'Mesh_size' - mesh size (knot to knot) (mm)
Column 2: 'y' - Smaller of the angels between mesh threads

```{r}
net.properties <- read.csv("~/kinneret modeling/selectivity/R code Feodor paper/Lobyerev-selectivity/Data/net properties.csv")
#Show table:
print(net.properties)
```

####Step 1.3: Add angel in Radian
```{r}
net.properties$Radian=0.018*net.properties$y
#Show table:
print(net.properties)
```

####Step 1.4: Define the angel between the end of upper and lower jaws (phi)
```{r}
phi_deg=7
phi_radian=0.018*phi_deg
```

####Step 1.5: Calculate the jaw length approximated by the linear function
```{r}
slope_jaw=1.02
intecept_jaw=3.41
Catch.data.Cod$Jaw_length=intecept_jaw + (slope_jaw*Catch.data.Cod$Length_group)
#Calculate h
Catch.data.Cod$h=Catch.data.Cod$Jaw_length * sin(phi_radian) * 2
print(head(Catch.data.Cod))
```
####Step 1.6: Merge the catch data with the net data
```{r}
Catch.data.Cod=merge(Catch.data.Cod,net.properties,by="Mesh_size")
#print first 6 lines
print(head(Catch.data.Cod))
```

###Step 2: Calculate P(O|C) and P(Th|C)
For calculating P(O|C) use eq. 4 in the paper

```{r}
#Calculate sin,cos,tan
Catch.data.Cod$sin=sin(0.5*Catch.data.Cod$Radian)
Catch.data.Cod$cos=cos(0.5*Catch.data.Cod$Radian)
Catch.data.Cod$tan=tan(0.5*Catch.data.Cod$Radian)
#Calculate P(O|C) Eq. 4
Catch.data.Cod$POC=(((Catch.data.Cod$cos * Catch.data.Cod$Mesh_size)-Catch.data.Cod$h)*((Catch.data.Cod$sin*Catch.data.Cod$Mesh_size)-(Catch.data.Cod$tan*Catch.data.Cod$h)))/(Catch.data.Cod$cos*Catch.data.Cod$sin*(Catch.data.Cod$Mesh_size)^2)
#Calculate P(Th|c)
Catch.data.Cod$PThC=1-Catch.data.Cod$POC
#Print first 6 rows
print(head(Catch.data.Cod))
```

###Step 3: Calculate P(W|E)

Import data on maximal girth (G_max). In final version girth information will be part of the complete catch data
```{r}
G_max_wide <- read.csv("~/kinneret modeling/selectivity/R code Feodor paper/Lobyerev-selectivity/Data/G_max_wide.csv", stringsAsFactors=FALSE)
#Transform to long format
G_max_long=gather(G_max_wide, t, G_max, m1:m73, factor_key=TRUE)
#Arrange table
#Drop NA
G_max_long=G_max_long[!is.na(G_max_long$length_group),]
G_max_long=G_max_long[!is.na(G_max_long$G_max),]
#Drop the second column 't'
G_max_long=G_max_long[ , !names(G_max_long) %in% c("t")]
#Order by first column
G_max_long=G_max_long[order(G_max_long$length_group),]
```

For each net calculate P(W|E)
```{r}
#Calculate the girth which each net will retain
net.properties$girth_retained=round(1.14*4*net.properties$Mesh_size)
#print(net.properties)
#Calculate the P(W|E) for each length group. P(W|E) is the ratio between fish over net.properties$girth_retained and the total number of fish in each group_size for each net type
PWE_t=expand.grid(net.properties$Mesh_size,unique(G_max_long$length_group)) %>% 
   dplyr::rename(Mesh_size = Var1,
          length_group = Var2)  %>%
   left_join(net.properties) %>% 
   left_join(G_max_long) %>% 
   select(Mesh_size,girth_retained,length_group,G_max) %>% 
   mutate(is.bigger = G_max >= girth_retained) %>% 
   group_by(Mesh_size,length_group) %>% 
   summarize(PWE = sum(is.bigger)/n()) %>%
  dplyr::rename(Length_group=length_group)
PWE_t=data.frame(PWE_t)
print(head(PWE_t))
Catch.data.Cod=merge(Catch.data.Cod,PWE_t ,by=c("Mesh_size",    "Length_group"))
```
###Step 4: Calculate Eq. 2

####Step 4.1: Calculate P(E|O) by the linear equation

```{r}
##Get the min and max size groups for each net
#Create table
net_PEO=data.frame(Mesh_size=unique(Catch.data.Cod$Mesh_size))
#Length group interval
Length_group_interval=2
#The minimal size for each net
min_wedged=data.frame(Catch.data.Cod %>%
   filter(!is.na(Wedged)) %>%
  group_by(Mesh_size) %>%
  summarize(min_size = min(Length_group, na.rm = TRUE)))
#The maximal size for each net
max_wedged=data.frame(Catch.data.Cod %>%
   filter(!is.na(Wedged)) %>%
  group_by(Mesh_size) %>%
  summarize(max_size = max(Length_group, na.rm = TRUE)))
#merge
net_PEO=merge(net_PEO,min_wedged,by="Mesh_size")
net_PEO=merge(net_PEO,max_wedged,by="Mesh_size")
#Substract and add the Length group interval
net_PEO$min_size=net_PEO$min_size-(Length_group_interval/2)
net_PEO$max_size=net_PEO$max_size+(Length_group_interval/2)
#Calculate the linear function
net_PEO$slope=-1/(net_PEO$max_size-net_PEO$min_size)
net_PEO$intercept=1-(net_PEO$slope*net_PEO$min_size)
#print
print(net_PEO)
#Calculate PEO
#Merge slope and intercept to data
Catch.data.Cod=merge(Catch.data.Cod,net_PEO,by="Mesh_size")
Catch.data.Cod$PEO=Catch.data.Cod$intercept+(Catch.data.Cod$Length_group)*Catch.data.Cod$slope
#Remove PEO values if no fish were wedged
Catch.data.Cod[is.na(Catch.data.Cod$Wedged),"PEO"]=NA
#Print first 6 lines
print(head(Catch.data.Cod))
```

####Step 4.2: Claculate Eq.2
```{r}
Catch.data.Cod$Ntotal=Catch.data.Cod$Wedged/(Catch.data.Cod$POC*Catch.data.Cod$PWE*Catch.data.Cod$PEO)
#Print first 6 lines
print(head(Catch.data.Cod))
```

###Step 5: Calculate CPUE
(table 21)
```{r}
#Define the number of trials
n_trials=11
#Aggregate catch for each mesh size
Catch.data.Cod$CPUE=rowSums(Catch.data.Cod[,c("Wedged","Tangled")],na.rm=T)/n_trials
#Print first 6 lines
print(head(Catch.data.Cod))
```

###Step 6: Calculate Nw per hour
(table 23)
```{r}
#Define number of hours of single (??) field trial
n_hours=12
Catch.data.Cod$Ntotal_per_hour=Catch.data.Cod$Ntotal/n_hours
#Print first 6 lines
print(head(Catch.data.Cod))
```

###Step 7: 
####Step 7.1: SLl,t for each length class as Ntotal-Qt
(table 24)
```{r}
Catch.data.Cod$SL_l_t=Catch.data.Cod$Ntotal_per_hour-Catch.data.Cod$CPUE
#Print first 6 lines
print(head(Catch.data.Cod))
```

####Step 7.2: N_AP calculation
First the table of Nlim (table 22c) is defined, in reality it is in input

!!!Notice- value of tau is very sensitive to the sum of the CPUE
```{r}
###The next data frame will be an experimental input
n_lim=data.frame(Mesh_size=c(20,25,30),Nlim=c(36,32,12))
#Sum CPUE per net
CPUE=data.frame(Catch.data.Cod %>%
  group_by(Mesh_size) %>%
  summarize(CPUE_sum = sum(CPUE, na.rm = TRUE)))
#merge to n_lim table
n_lim=merge(n_lim,CPUE,by="Mesh_size")
#Calculate tau (table 25)
n_lim$tau=n_hours/(-log(1-(n_lim$CPUE/n_lim$Nlim)))
n_lim$N_AP=((n_lim$Nlim*(1.71))/n_lim$tau)*n_hours
print(n_lim)
```

####Step 7.3: Size specific N_AP
!!!In table 26 you use data from table 24 (SL)- is SL=0 if there is no value for this length group in table 24?
```{r}
#Replace NA's in column 'SL_l_t' with 0
Catch.data.Cod[is.na(Catch.data.Cod$SL_l_t),"SL_l_t"]=0
#merge 
Catch.data.Cod=merge(Catch.data.Cod,n_lim,by="Mesh_size")
#Calculate N_AP per size class
Catch.data.Cod$N_AP_size=(Catch.data.Cod$N_AP*(Catch.data.Cod$CPUE/Catch.data.Cod$CPUE_sum))+Catch.data.Cod$SL_l_t
print(head(Catch.data.Cod))
```

###Step 8: Calculate selectivity
(table 28)
```{r}
##Calculate selectivity
Catch.data.Cod$selectivity=Catch.data.Cod$CPUE/Catch.data.Cod$Ntotal_per_hour
##plot
Catch.data.Cod$Mesh_size_fac=as.character(Catch.data.Cod$Mesh_size)
ggplot(Catch.data.Cod, aes(x=Length_group, y=selectivity, group=Mesh_size_fac)) +
  geom_line(aes(color=Mesh_size_fac))+
  geom_point(aes(color=Mesh_size_fac))+
   labs(x = "length Group",y="Selectivity",color="Mesh Size")
  
