---
title: "Package 'Netselect'"
author: "Itai van Rijn"
date: "March 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)
```
###General package information (will be needed later for publishing the package)
Version XXX

Date XXX

Title Gill net selectivity model

Author XXX

Maintainer XXX <mail@adress>

Depends (R packages required)

Description Functions for calculating gill net selectivity following Lobyerev & Hoffman 
(2018) Selectivity model 

License GPL (>= 2)

RoxygenNote XXX

NeedsCompilation no

Repository CRAN

Date/Publication XXX

###List of functions

R topics documented:

###Description of data and functions

***

#####__'Cod_selectivity_data'__ (_data_)- Field survey data for calculating selectivity for multiple mesh-size

__Description__

The Cod_selectivity_data data frame has 694 rows and 6 columns. This field data was used for calculation of net selectivity for Cod (_Gadus morhua_)

__Usage__

Cod_selectivity_data

__Format__

This data frame contains the following columns:

__Survey_index__ - Unique ID for each field trial

__Net_name__ - Name of the net used

__Mesh_size__ - Mesh size (knot to knot) (mm)

__Length_FL__ - Fork length (cm)

__Maximal_girth__ - Maximal body girth (mm)

__Catch_method__ - Tangled or Wedged

__Source__

(Add some Ref. to the data)

```{r}
Cod_selectivity_data <- read.csv("~/kinneret modeling/selectivity/R code Feodor paper/Lobyerev-selectivity/Data/Cod_selectivity_data.csv")
print(head(Cod_selectivity_data ))
```

***

***

#####__'Net_properties_data'__ (_data_)- Gill nets properties for calculating selectivity
__Description__

The Cod_selectivity_data data frame has 694 rows and 6 columns. This field data was used for calculation of net selectivity for Cod (_Gadus morhua_)

__Usage__

Net_properties_data

__Format__

This data frame contains the following columns:

__Mesh_size__ - Mesh size (knot to knot) (mm)

__y__ - Smaller of the angels between mesh threads (deg)

__Source__

(Add some Ref. to the data)

```{r}
Net_properties_data <- read.csv("~/kinneret modeling/selectivity/R code Feodor paper/Lobyerev-selectivity/Data/net properties.csv")
#Show table:
print(Net_properties_data)
```

***

***

#####__'Mouth_dimension'__ (_function_)- Fish mouth openning dimension

__Description__

Calculate the height (h) of swimming fish mouth openning  

__Usage__

Mouth_dimension(Survey_data = NULL, Net_data = NULL, slope_jaw = NULL, intercept_jaw = NULL, phi = NULL, length_group_bin=2)

__Arguments__

Survey_data - data frame of field survey data containing information on body length (see: 'Cod_selectivity_data')

Net_data - data frame of gill net properties (see: 'Net_properties_data') 

slope_jaw - numeric value of the slope term of the linear equation concting body length to jaw length

intercept_jaw - numeric value of the intercept term of the linear equation concting body length to jaw length

phi - numeric angle between the ends of upper and lower jaws when the mouth is opened during ram ventilation

length_group_bin - numeric value of the bin size used to aggregate fish to length groups

__Details__

__Value__

Data frame similar to arrgument 'Survey_data' with additional columns

__Author(s)__

XXX

__Function structure__

```{r}
Mouth_dimension=function(Survey_data,Net_data,slope_jaw,intercept_jaw,phi,length_group_bin=2){
  ##Add angle in radians to Net_data
  Net_data$Radian=0.018*Net_data$y
  ##Transform phi to eadians
  phi_radian=0.018*phi
  ##Calculate Jaw length from linear function
  Survey_data$Jaw_length=intercept_jaw + (slope_jaw*Survey_data$Length_FL)
  ##Calculate height of mouth openning
  Survey_data$h=Survey_data$Jaw_length * sin(phi_radian) * 2
  return(Survey_data)
}
```

__Examples__

```{r}
Mouth_dimension_output=Mouth_dimension(Survey_data=Cod_selectivity_data,
                                       Net_data=Net_properties_data,
                                       slope_jaw=1.02,
                                       intercept_jaw=3.41,
                                       phi=7)

print(head(Mouth_dimension_output))
```

***

***

#####__'POC_PThC'__ (_function_)- Probabilities of contacting gill net at openning or thread

__Description__

Calculate P(O|C) and P(Th|C). 

__Usage__

POC_PThC(Survey_and_mouth_data = NULL, Net_data = NULL, length_group_bin=2)

__Arguments__

Survey_and_mouth_data - data frame of field survey data and mouth openning dimension (an output of function 'Mouth_dimension')

Net_data - data frame of gill net properties (see: 'Net_properties_data') 

length_group_bin - numeric value of the bin size used to aggregate fish to length groups

__Details__

__Value__

Data frame similar to arrgument 'Survey_mouth_data' with additional columns: POC and PThC

__Author(s)__

XXX

__Function structure__

```{r}
POC_PthC=function(Survey_and_mouth_data,Net_data,length_group_bin=2){
  #Create length groups based on fish FL
  Survey_data$Length_group=length_group_bin*round(Survey_and_mouth_data$Length_FL/length_group_bin)
  ##
  Survey_data=Survey_data[unique(Survey_data[,c("Mesh_zise","Length_group")]),]
  ##Add angle in radians to Net_data
  Net_data$Radian=0.018*Net_data$y
  ##Merge gill net properties table to survey data
  Survey_and_mouth_data=merge(Survey_and_mouth_data,Net_data,by="Mesh_size")
  ##Calculate sin,cos,tan
  Survey_and_mouth_data$sin=sin(0.5*Survey_and_mouth_data$Radian)
  Survey_and_mouth_data$cos=cos(0.5*Survey_and_mouth_data$Radian)
  Survey_and_mouth_data$tan=tan(0.5*Survey_and_mouth_data$Radian)
  #Calculate P(O|C) Eq. 4
  Survey_and_mouth_data$POC=(((Survey_and_mouth_data$cos * Survey_and_mouth_data$Mesh_size)-Survey_and_mouth_data$h)*((Survey_and_mouth_data$sin*Survey_and_mouth_data$Mesh_size)-(Survey_and_mouth_data$tan*Survey_and_mouth_data$h)))/(Survey_and_mouth_data$cos*Survey_and_mouth_data$sin*(Survey_and_mouth_data$Mesh_size)^2)
  #Calculate P(Th|c)
  Survey_and_mouth_data$PThC=1-Survey_and_mouth_data$POC
  Survey_and_mouth_data=Survey_and_mouth_data[unique(Survey_and_mouth_data[,c("Mesh_size","Length_group")])]
  return(Survey_and_mouth_data)
}
```

__Examples__

```{r}
##Run function '
POC_PthC_output=POC_PthC(Survey_and_mouth_data=Mouth_dimension_output,
                                       Net_data=Net_properties_data)


print(head(POC_PthC_output))
```

***

***

#####__'PEO'__ (_function_)- Probabilities of fish entering net openning if contacted an openning

__Description__

Calculate P(E|O). 

__Usage__

PEO(Survey_data = NULL, Net_data = NULL, length_group_bin=2)

__Arguments__

Survey_data - data frame of field survey data containing information on body length (see: 'Cod_selectivity_data')

Net_data - data frame of gill net properties (see: 'Net_properties_data')

length_group_bin - numeric value of the bin size used to aggregate fish to length groups

__Details__

__Value__

Data frame similar to arrgument 'Survey_data' with additional columns: PWE

__Author(s)__

XXX

__Function structure__

```{r}
PEO=function(Survey_data,Net_data,length_group_bin=2){
  #Create length groups based on fish FL
  Survey_data$Length_group=length_group_bin*round(Survey_data$Length_FL/length_group_bin)
  #The minimal and maximal size of wedged fish in net type
  net_PEO=data.frame(Survey_data %>%
                       filter(Catch_method=="Wedged") %>%
                       group_by(Mesh_size) %>%
                       summarize(min_FL = min(Length_group, na.rm = TRUE),
                                 max_FL = max(Length_group, na.rm = TRUE))%>%
                       left_join(Net_data))
  #Substract and add the Length group interval
  net_PEO$min_FL=net_PEO$min_FL-(length_group_bin/2)
  net_PEO$max_FL=net_PEO$max_FL+(length_group_bin/2)
  #Calculate the linear function
  net_PEO$slope=-1/(net_PEO$max_FL-net_PEO$min_FL)
  net_PEO$intercept=1-(net_PEO$slope*net_PEO$min_FL)
  #Calculate PEO
  #Merge slope and intercept to data
  Survey_data=merge(Survey_data,net_PEO,by="Mesh_size")
  Survey_data$PEO=Survey_data$intercept+(Survey_data$Length_group)*Survey_data$slope
  #Remove negative PEO
  Survey_data[Survey_data$PEO<0,"PEO"]=NA
  #If PEO>1 than the probability is 1
  Survey_data$PEO = ifelse(Survey_data$PEO > 1, 1,Survey_data$PEO)
  #Table of PEO for unique Mesh_size and Length_group
  output_table=Survey_data[!duplicated(Survey_data[,c("Mesh_size","Length_group")]),]
  return(output_table)
}
```

__Examples__

```{r}
##Run function '
PEO_output=PEO(Survey_data=Cod_selectivity_data,Net_data=Net_properties_data,length_group_bin=2)


print(head(PEO_output))
```

***

***

#####__'PWE'__ (_function_)- Probabilities of fish being wedged if entered net openning

__Description__

Calculate P(W|E). 

__Usage__

PWE(Survey_data = NULL, Net_data = NULL, length_group_bin=2)

__Arguments__

Survey_data - data frame of field survey data containing information on body length (see: 'Cod_selectivity_data')

Net_data - data frame of gill net properties (see: 'Net_properties_data')

length_group_bin - numeric value of the bin size used to aggregate fish to length groups

__Details__

__Value__

Data frame similar to arrgument 'Survey_data' with additional columns: PWE

__Author(s)__

XXX

__Function structure__

```{r}
PWE=function(Survey_data,Net_data,length_group_bin=2){
  ##Calculate the girth retained in mesh size
 Net_data$girth_retained=round(1.14*4*Net_data$Mesh_size)
 #Create length groups based on fish FL
 Survey_data$Length_group=length_group_bin*round(Survey_data$Length_FL/length_group_bin) 
  #Calculate the P(W|E) for each length group. 
  PWE=expand.grid(Net_data$Mesh_size,unique(Survey_data$Length_group)) %>% 
    dplyr::rename(Mesh_size = Var1,
           Length_group = Var2)  %>%
    left_join(Net_data) %>% 
    left_join(Survey_data) %>% 
    select(Mesh_size,girth_retained,Length_group,Maximal_girth) %>% 
    mutate(is.bigger = Maximal_girth >= girth_retained) %>% 
    group_by(Mesh_size,Length_group) %>% 
    summarize(PWE = sum(is.bigger)/n()) 
  PWE=data.frame(PWE)
  return(PWE)
}
```

__Examples__

```{r}
##Run function '
PWE_output=PWE(Survey_data=Cod_selectivity_data,Net_data=Net_properties_data,length_group_bin=2)


print(head(PWE_output))
```

***

***

#####__'Ntotal'__ (_function_)- Number of fish which contacted the net

__Description__

Calculate the number of fish which contacted the net. This function contains other functions from this package: 'Mouth_dimension','POC_PthC','PEO' and 'PWE'

__Usage__

Ntotal(Survey_data = NULL, Net_data = NULL,slope_jaw = NULL, intercept_jaw = NULL, phi = NULL, length_group_bin=2)

__Arguments__

Survey_data - data frame of field survey data containing information on body length (see: 'Cod_selectivity_data')

Net_data - data frame of gill net properties (see: 'Net_properties_data')

slope_jaw - numeric value of the slope term of the linear equation concting body length to jaw length

intercept_jaw - numeric value of the intercept term of the linear equation concting body length to jaw length

phi - numeric angle between the ends of upper and lower jaws when the mouth is opened during ram ventilation

length_group_bin - numeric value of the bin size used to aggregate fish to length groups

__Details__

__Value__



__Author(s)__

XXX

__Function structure__

```{r}
Ntotal=function(Survey_data,Net_data,slope_jaw,intercept_jaw,phi,length_group_bin=2){
  ##Apply function 'Mouth_dimension'
  Mouth_dimension_output=Mouth_dimension(Survey_data=Survey_data,
                                       Net_data=Net_data,
                                       slope_jaw=slope_jaw,
                                       intercept_jaw=intercept_jaw,
                                       phi=phi)
  ##Apply function 'POC_PthC'
  POC_PthC_output=POC_PthC(Survey_and_mouth_data=Mouth_dimension_output,
                                       Net_data=Net_data)
   ##Apply function 'PEO'
  PEO_output=PEO(Survey_data=POC_PthC_output,Net_data=Net_properties_data,length_group_bin=2)
     ##Apply function 'PWE'
  PWE_output=PWE(Survey_data=Survey_data,Net_data=Net_data,length_group_bin=length_group_bin)
  ##Merge 'PEO_output' with 'PWE_output'
  Complete_data=merge(PEO_output,PWE_output,by=c("Mesh_size","Length_group"))
  #Aggrigate data by net type and length group
  aggrigated_data=Complete_data %>%
    group_by(Mesh_size, Length_group,Catch_method,POC,PEO,PWE)
  aggrigated_data <- summarise(aggrigated_data,count = n())
  #Subset to wedged individuals
  aggrigated_data=aggrigated_data[aggrigated_data$Catch_method=="Wedged",]
  #Calculate the number of fish which contacted the net
aggrigated_data$Ntotal=aggrigated_data$count/(aggrigated_data$POC*aggrigated_data$PWE*aggrigated_data$PEO)
return(aggrigated_data)}
```

__Examples__

```{r}
##Run function '
PWE_output=PWE(Survey_data=Cod_selectivity_data,Net_data=Net_properties_data,length_group_bin=2)


print(head(PWE_output))
```

***
