---
title: "Package 'Netselect'"
author: "Itai van Rijn"
date: "March 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###General package information (will be needed later for publishing the package)
Version XXX

Date XXX

Title Gill net selectivity model

Author XXX

Maintainer XXX <mail@adress>

Depends (R packages required)

Description Functions for calculating gill net selectivity following Lobyerev & Hoffman 
(2018) Selectivity model 

License GPL (>= 2)

RoxygenNote XXX

NeedsCompilation no

Repository CRAN

Date/Publication XXX

###List of functions

R topics documented:

###Description of data and functions

#####__'Cod_selectivity_data'__ - Field survey data for calculating selectivity for multiple mesh-size

__Description__

The Cod_selectivity_data data frame has 694 rows and 6 columns. This field data was used for calculation of net selectivity for Cod (_Gadus morhua_)

__Usage__

Cod_selectivity_data

__Format__

This data frame contains the following columns:

__Survey_index__ - Unique ID for each field trial

__Net_name__ - Name of the net used

__Mesh_size__ - Mesh size (mm knot-to-knot)

__Length_FL__ - Fork length (cm)

__Maximal_girth__ - Maximal body girth (mm)

__Catch_method__ - Tangled or Wedged

__Source__

(Add some Ref. to the data)

```{r}
Cod_selectivity_data <- read.csv("~/kinneret modeling/selectivity/R code Feodor paper/Lobyerev-selectivity/Data/Cod_selectivity_data.csv")
print(head(Cod_selectivity_data ))
```


